\vspace{-3em} \textbf{ Mapping stellar surfaces\\ II: An interpretable Gaussian process model for light curves }

### 2021-01-29 paper2/ms.tex
minor tweaks

Fortunately, there is a better way to reduce the number of parameters in
the problem: we can explicitly marginalize over the stellar inclination.
That is, we may write the mean and covariance of our GP as
%
%
\begin{align}
    \label{eq:mu_marg}
    \pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet)
     & = \mathrm{E} \Big[ \mathbbb{f} \, \Big| \, P, \mathbf{u}, \pmb{\theta}_\bullet \Big]
    \nonumber                                                                                                                                                                                                        \\
     & = \mathbf{1} + \mathbf{e}_I
    \\[0.5em]
    \label{eq:cov_marg}
    \pmb{\Sigma}(P, \mathbf{u}, \pmb{\theta}_\bullet)
     & = \mathrm{E} \Big[ \mathbbb{f} \, \mathbbb{f}^\top \, \Big| \, P, \mathbf{u}, \pmb{\theta}_\bullet \Big] - \pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet) \pmb{\mu}^\top(P, \mathbf{u}, \pmb{\theta}_\bullet)
    \nonumber                                                                                                                                                                                                        \\
     & =
    \mathbf{E}_I
    -
    \mathbf{e}_I \,
    \mathbf{e}_I^\top
\end{align}
%
where we define the inclination first moment integral
%
\begin{align}
    \label{eq:eI}
    \mathbf{e}_I
     & \equiv
    \int
    \pmb{\mathcal{A}}(\mathbb{I}, P, \mathbf{u}) \,
    \mathrm{E} \Big[ \mathbbb{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
    p(\mathbb{I}) \, \mathrm{d}\mathbb{I}
\end{align}
%
and the inclination second moment integral
%
\begin{align}
    \label{eq:EI}
    \mathbf{E}_I
     & \equiv
    \int
    \pmb{\mathcal{A}}(\mathbb{I}, P, \mathbf{u}) \,
    \mathrm{E} \Big[ \mathbbb{y} \, \mathbbb{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] \,
    \pmb{\mathcal{A}}^\top(\mathbb{I}, P, \mathbf{u})
    p(\mathbb{I}) \, \mathrm{d}\mathbb{I}
    \quad.
\end{align}
%
The expectations inside the integrals in the expressions for
$\mathbf{e}_I$ and $\mathbf{E}_I$
are given by
Equations~(\ref{eq:exp_y}) and (\ref{eq:exp_yy}), respectively, and
are computed in Appendix~\ref{sec:integrals}.
%
If we are able to perform the integrals in those expressions,
we can \emph{dramatically} reduce the number of
parameters in our ensemble problem.
%
As we show in Appendix~\ref{sec:gp-inc}, if we assume that stellar
inclinations are distributed isotropically, these integrals
do in fact have closed-form solutions.

@@

But before we dive in, it is worth pausing for a moment to take another
look at Figure~\ref{fig:samples}. While we focused on the shortcomings of
the GP as a prior in the discussion above, it is important to appreciate
that it even works in the first place!
%Despite some issues, the fact that
%the surface map samples \emph{actually look like maps of spotted stellar surfaces}
%is fairly amazing.
A Gaussian process, after all, is a non-parameteric
\emph{process} describing a smooth and continuous function only via its
covariance structure. The GP knows nothing about the existence of discrete
spots---only how any two points on the surface are correlated.
Because spherical harmonics are smooth functions with support over the
entire sphere, the GP also does not know about features restricted
to certain latitudes; in fact, in most applications of GPs to mapping
problems in astronomy \citep[such as in models of the cosmic microwave
    background; e.g.,][]{Wandelt2012}, the process is assumed to be isotropic,
with no preferred angular direction.
The fact that we are able to construct a GP that captures discrete
features restricted to certain regions on the surface (albeit with
some issues) is remarkable.

>>>>

Fortunately, there is a better way to reduce the number of parameters in
the problem: we can explicitly marginalize over the stellar inclination.
That is, we may write the mean and covariance of our GP as
%
%
\begin{align}
    \label{eq:mu_marg}
    \pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet)
     & = \mathrm{E} \Big[ \mathbbb{f} \, \Big| \, P, \mathbf{u}, \pmb{\theta}_\bullet \Big]
    \nonumber                                                                                                                                                                                                        \\
     & = \mathbf{1} + \mathbf{e}_I
    \\[0.5em]
    \label{eq:cov_marg}
    \pmb{\Sigma}(P, \mathbf{u}, \pmb{\theta}_\bullet)
     & = \mathrm{E} \Big[ \mathbbb{f} \, \mathbbb{f}^\top \, \Big| \, P, \mathbf{u}, \pmb{\theta}_\bullet \Big] - \pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet) \pmb{\mu}^\top(P, \mathbf{u}, \pmb{\theta}_\bullet)
    \nonumber                                                                                                                                                                                                        \\
     & =
    \mathbf{E}_I
    -
    \mathbf{e}_I \,
    \mathbf{e}_I^\top
\end{align}
%
where we define the inclination first moment integral
%
\begin{align}
    \label{eq:eI}
    \mathbf{e}_I
     & \equiv
    \int
    \pmb{\mathcal{A}}(\mathbb{I}, P, \mathbf{u}) \,
    \mathrm{E} \Big[ \mathbbb{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
    p(\mathbb{I}) \, \mathrm{d}\mathbb{I}
\end{align}
%
and the inclination second moment integral
%
\begin{align}
    \label{eq:EI}
    \mathbf{E}_I
     & \equiv
    \int
    \pmb{\mathcal{A}}(\mathbb{I}, P, \mathbf{u}) \,
    \mathrm{E} \Big[ \mathbbb{y} \, \mathbbb{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] \,
    \pmb{\mathcal{A}}^\top(\mathbb{I}, P, \mathbf{u})
    p(\mathbb{I}) \, \mathrm{d}\mathbb{I}
    \quad,
\end{align}
%
and $\mathbb{I}$ is the random variable corresponding to the inclination.
%
The expectations inside the integrals in the expressions for
$\mathbf{e}_I$ and $\mathbf{E}_I$
are given by
Equations~(\ref{eq:exp_y}) and (\ref{eq:exp_yy}), respectively, and
are computed in Appendix~\ref{sec:integrals}.
%
If we are able to perform the integrals in those expressions,
we can \emph{dramatically} reduce the number of
parameters in our ensemble problem.
%
As we show in Appendix~\ref{sec:gp-inc}, if we assume that stellar
inclinations are distributed isotropically, these integrals
do in fact have closed-form solutions.

@@

But before we dive into calibration tests, it is worth pausing for a moment to take another
look at Figure~\ref{fig:samples}. While we focused on the shortcomings of
the GP as a prior in the discussion above, it is important to appreciate
that it even works in the first place!
%Despite some issues, the fact that
%the surface map samples \emph{actually look like maps of spotted stellar surfaces}
%is fairly amazing.
A Gaussian process, after all, is a non-parameteric
\emph{process} describing a smooth and continuous function only via its
covariance structure. The GP knows nothing about the existence of discrete
spots---only how any two points on the surface are correlated.
Because spherical harmonics are smooth functions with support over the
entire sphere, the GP also does not know about features restricted
to certain latitudes; in fact, in most applications of GPs to mapping
problems in astronomy \citep[such as in models of the cosmic microwave
    background; e.g.,][]{Wandelt2012}, the process is assumed to be isotropic,
with no preferred angular direction.
However, by prescribing the correct structure to the covariance matrix,
we are able to approximately model compact spot-like
features with given sizes and restricted to particular latitudes.
